---
title: "combine_responses"
author: "Wulcan"
date: "2024-08-17"
output: html_document
---

# Objective
Combine human and llm responses

```{r libraries, directories and functions}
library(tidyverse)

toBox <- "~/Library/CloudStorage/Box-Box/"

intermediatesDir <-paste0(toBox,"Projects/LLM/llm_vet_records/intermediates/combine_responses/")

combine_responses <- function(response_parsed, llm_parsed) {
  # Identify columns in response_parsed but not in llm_parsed
  missing_in_llm <- setdiff(names(response_parsed), names(llm_parsed))
  
  # Identify columns in llm_parsed but not in response_parsed
  missing_in_response <- setdiff(names(llm_parsed), names(response_parsed))
  
  # Add missing columns to llm_parsed with NA values
  if(length(missing_in_llm) > 0) {
    for(col in missing_in_llm) {
      llm_parsed[[col]] <- NA
    }
  }
  
  # Add missing columns to response_parsed with NA values
  if(length(missing_in_response) > 0) {
    for(col in missing_in_response) {
      response_parsed[[col]] <- NA
    }
  }
  
  # Combine the two dataframes using bind_rows
  response <- dplyr::bind_rows(response_parsed, llm_parsed)
  
  return(response)
}

```

## Combine human and GPT4o response for test set
### Input
- test_response_parsed (intermediates/parse_survey)
- test_GPT4o_parsed (intermediates/parse_llm_output)

## Output
- test_response_human_GPT4o (intermediates/combine responses)

```{r test GPT4o}
test_response_parsed <- readRDS(paste0(toBox, "Projects/LLM/llm_vet_records/intermediates/parse_survey/test_response_parsed.rds"))

test_GPT4o_parsed <- readRDS(paste0(toBox, "Projects/LLM/llm_vet_records/intermediates/parse_llm_output/test_GPT4o_corrected_parsed.rds"))

test_response_human_GPT4o <- combine_responses(test_response_parsed, test_GPT4o_parsed)

saveRDS(test_response_human_GPT4o, paste0(intermediatesDir, "test_response_human_GPT4o.rds"))
```

## Combine human and GPT35 response for test set
### Input
- test_response_parsed (intermediates/parse_survey)
- test_GPT35_parsed (intermediates/parse_llm_output)

## Output
- test_response_human_GPT35 (intermediates/combine responses)

```{r test GPT4o}
test_response_parsed <- readRDS(paste0(toBox, "Projects/LLM/llm_vet_records/intermediates/parse_survey/test_response_parsed.rds"))

test_GPT35_parsed <- readRDS(paste0(toBox, "Projects/LLM/llm_vet_records/intermediates/parse_llm_output/test_GPT35_parsed.rds"))

test_response_human_GPT35 <- combine_responses(test_response_parsed, test_GPT35_parsed)

saveRDS(test_response_human_GPT35, paste0(intermediatesDir, "test_response_human_GPT35.rds"))
```