---
title: "estimate_prevalence"
author: "Wulcan"
date: "2024-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, directories and functions}
library(tidyverse)
library(readxl)

toBox <- "~/Library/CloudStorage/Box-Box/"
#toBox <- "C:/Users/15303/Box/"

inputDir <- paste0(toBox,"Projects/LLM/llm_vet_records/input/")

intermediatesDir <- paste0(toBox,"Projects/LLM/llm_vet_records/intermediates/estimate_prevalence/")

outputDir <- paste0(toBox,"Projects/LLM/llm_vet_records/output/estimate_prevalence/")

# Function to detect mentions of specified words in specified variables

detect_mentions <- function(df, variables, words) {
  
  # Iterate over each word to generate the detection columns
  for (word in words) {
    # Define the function to check for the presence of the word
    check_function <- function(text) {
      str_detect(text, regex(paste0("\\b", word, "\\w*"), ignore_case = TRUE))
    }
    
    # Generate the column names dynamically
    mention_col <- gsub(" ", "_", word)
    mention_field_col <- paste0("mentions_", word, "_in_field")
    
    # Apply the detection logic
    df <- df %>%
      mutate(
        !!mention_col := pmap_lgl(select(., all_of(variables)), 
                                  ~ any(map_lgl(list(...), check_function))),
        !!mention_field_col := pmap(select(., all_of(variables)), 
                                    ~ paste(variables[map_lgl(list(...), check_function)], collapse = ", "))
      ) %>%
      # Replace empty strings with NA in the "_in_field" columns
      mutate(!!mention_field_col := ifelse(.data[[mention_field_col]] == "", NA, .data[[mention_field_col]]))
  }

  return(df)
}

# Function to calculate prevalence statistics
calculate_prevalence <- function(df, mention_cols) {
  
  # Calculate the total number of observations
  total_observations <- nrow(df)
  
  # Calculate the percentage of records where each mentions_* column is TRUE
  percentage_of_records <- sapply(mention_cols, function(col) {
    true_count <- sum(df[[col]], na.rm = TRUE)
    percentage <- (true_count / total_observations) * 100
    return(percentage)
  })
  
  # Create the prevalence data frame
  prevalence_df <- data.frame(
    clinical_sign =  mention_cols,
    percentage_of_records = as.numeric(percentage_of_records),
    total_observations = total_observations
  )
  
  rownames(prevalence_df) <- NULL
  
  return(prevalence_df)
}

```

## Estimate prevalence 

### Objective
Estimate prevalence of GI signs in EHRs by different methods (key-word search and manual review) of different sets of EHRs (full set and pilot samples) using different EHR fields (full vs history). 

### Input
- contains_free_text.rds <VMACS/all_cats_ever/output/contains_free_text>
- pilot_free_text_graded.xlsx (manually graded from VMACS) 
- contains_hx <VMACS/all_cats_ever/output/contains_free_text>
- pilot_hx_response_parsed (intermediates/parse_survey)

### Output
- key_gi_free_text (intermediates/estimate_prevalence)
- prevalence_gi (output/estimate_prevalence)

### Issues/Deviations
- pilot_free_text_graded consists of manual review of complete VMACS EHR's in a google sheet (My Drive > FeGI > llm > tuning) and (My Drive > FeGI > llm > tuning_limit). Only record 1:36 of the pilot sample (pilot_free_text) were assessed and the extent of assessment was gradually decreased through the 36 EHR (1:3 presence, chronicity, duration ~ 15 minutes/EHR; 1:22 presence and chronicity ~3 min/EHR; 23:36 presence ~2 min/EHR). Two EHRs had insufficient info for grading. 
- The key word searches used "anorexia" and manual review "decreased appetite"

```{r estimate prevalence}

# Estimate prevalence by Key-word search of full database all EHR fields
contains_free_text <- readRDS(paste0(inputDir, "contains_free_text.rds"))

free_text_variables  <- c("presentingComplaint_visit", "pertinentHistory_visit", "physicalExamination_visit", "problems_visit",  "procedures_visit" , "plans_and_progress_notes_visit", "clinicalDx_visit" , "comments_visit", "dischargeSummary_visit", "dischargeInstructions_visit")

words <- c("vomit", "diarrhea", "polyphagia", "anorexia", "constipation", "weight loss")

# Detect mentions in the dataframe
key_gi_free_text <- detect_mentions(contains_free_text, free_text_variables, words)

saveRDS(key_gi_free_text, paste0(intermediatesDir, "key_gi_free_text.rds"))

# Estimate prevalence by Key-word search of full database all EHR fields
 words_modified <- gsub(" ", "_", words)
 
# Calculate prevalence
prevalence_key_full_all <- calculate_prevalence(key_gi_free_text, words_modified) %>%
  mutate(across(everything(), ~ replace(., . == "vomit", "vomiting"))) %>%
  mutate(across(everything(), ~ replace(., . == "anorexia", "decreased_appetite"))) %>%
  rename(percentage_by_key_word_full_database_all_fields = percentage_of_records) %>%
  rename(n_key_word_full_database_all_fields = total_observations)

#Estimate prevalence by manual review of EHRs in pilot set all EHR fields
manually_checked <- read_excel(paste0(inputDir, "pilot_free_text_graded.xlsx")) %>%
  mutate(across(where(is.character), ~ ifelse(. == "NA", NA, .))) %>%
  filter(!is.na(lethargy_presence)) %>%
  filter(!is.na(time_start)) # not graded

clinical_signs <- c("decreased_appetite", "polyphagia", "vomiting", "diarrhea", "constipation", "weight_loss")

variables <-  c("decreased_appetite_presence", "polyphagia_presence", "vomiting_presence", "diarrhea_presence", "constipation_presence", "weight_loss_presence")

# Convert character columns to logical
manually_checked <- manually_checked %>%
  mutate_at(vars(variables), as.logical)

prevalence_manual_pilot_all <- calculate_prevalence(manually_checked, variables) %>%
  mutate(clinical_sign = gsub("_presence", "", clinical_sign)) %>%
  rename(percentage_by_manual_review_pilot_all_fields = percentage_of_records) %>%
  rename(n_manual_review_pilot_all_fields = total_observations)

#Estimate prevalence by key word search of EHRs in full database, history field
contains_hx <- readRDS(paste0(inputDir, "contains_hx.rds"))

hx_variables <- c ("pertinentHistory_visit")

words <- c("vomit", "diarrhea", "polyphagia", "anorexia", "constipation", "weight loss")

# Detect mentions in the dataframe
key_gi_hx <- detect_mentions(contains_hx, hx_variables, words) 

#calculate prevalence
words_modified <- gsub(" ", "_", words)

prevalence_key_full_hx <- calculate_prevalence(key_gi_hx, words_modified) %>%
  mutate(across(everything(), ~ replace(., . == "vomit", "vomiting"))) %>%
  mutate(across(everything(), ~ replace(., . == "anorexia", "decreased_appetite"))) %>%
  rename(percentage_by_key_word_full_database_history = percentage_of_records) %>%
  rename(n_key_word_full_database_hx = total_observations)


#Estimate prevalence by manual review of EHRs in pilot set (pilot_hx), history field

response_parsed <- readRDS(paste0(toBox, "Projects/LLM/llm_vet_records/intermediates/parse_survey/pilot_hx_response_parsed.rds"))

variables <- c("vomiting", "diarrhea", "polyphagia", "decreased_appetite", "constipation", "weight_loss")

prevalence_manual_pilot_hx <- calculate_prevalence(response_parsed, variables)  %>%
  rename(percentage_by_manual_review_pilot_hx = percentage_of_records) %>%
  rename(n_key_word_manual_review_pilot_hx  = total_observations)

# Estimates prevalence by manual review of 250 records by 5 observers
response_mode_test_GPT4o <- readRDS(paste0(toBox,"Projects/LLM/llm_vet_records/intermediates/add_mode/mode_test_GPT4o.rds")) %>%
  filter(respondent_type == "Humans") %>%
  select(case, question, mode) %>%
  pivot_wider(names_from = question, values_from = mode)

variables <- c("vomiting", "diarrhea", "polyphagia", "decreased_appetite", "constipation", "weight_loss")

prevalence_manual_test_hx <- calculate_prevalence(response_mode_test_GPT4o, variables)  %>%
  rename(percentage_by_manual_review_test_hx = percentage_of_records) %>%
  rename(n_key_word_manual_review_test_hx  = total_observations)
  

# List of data frames to merge
df_list <- list(prevalence_key_full_all, prevalence_manual_pilot_all, prevalence_key_full_hx, prevalence_manual_pilot_hx, prevalence_manual_test_hx)

# Merge all data frames in the list by a common column, e.g., "clinical_sign"
prevalence <- Reduce(function(x, y) merge(x, y, by = "clinical_sign", all = TRUE), df_list)


write_csv(prevalence, paste0(outputDir, "prevalence_gi.csv"))
```



                                    


