---
title: "make_survey"
author: "Wulcan"
date: "2024-08-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objective
Generate a qualtrics survey from a set of records
```{r libraries, directories and functions}
library(tidyverse)
library(readxl)
library(jsonlite)

toBox <- "~/Library/CloudStorage/Box-Box/"
#toBox <- "C:/Users/15303/Box/" #lenovo

inputDir <- paste0(toBox,"Projects/LLM/llm_vet_records/input/")

intermediatesDir <- paste0(toBox,"Projects/LLM/llm_vet_records/intermediates/make_survey/")

# Function to escape HTML tags (to keep <> visible)
escape_html <- function(text) {
  text <- gsub("<", "&lt;", text)
  text <- gsub(">", "&gt;", text)
  return(text)
}

generate_survey_text <- function(data) {
  survey_text <- "[[AdvancedFormat]]\n\n"
  
  for (i in 1:nrow(data)) {
    case_no <- data$case[i]
    admission_date <- data$admission_date_visit[i]
    presenting_complaint <- data$presentingComplaint_visit[i]
    history <- data$pertinentHistory_visit[i]
    
     block_text <- paste0(
      "[[Block: Case ", case_no, "]]\n\n",
      
      "[[Question: Timing]]\n",
      "[[ID: case_", case_no, "_timing]]\n\n",
      
      "[[Question: DB]]\n",
      "[[ID: case_", case_no, "_history]]\n",
      "<b>Case number:</b>\n", case_no, "<br><br>",
      
      "<b>Admission date:</b>\n", admission_date, "<br><br>",
      "<b>Presenting complaint:</b>\n", presenting_complaint, "<br><br>",
      "<b>History:</b>\n", history,"\n\n",
      
      "[[Question: TE]]\n",
      "[[ID: case_", case_no, "_appetite_reference]]\n",
      "<b>Appetite:</b>\n<br><br>",
      "Reference: \n\n",
      
      "[[Question: MC:SingleAnswer]]\n",
      "[[ID: case_", case_no, "_decreased_appetite]]\n",
      "Decreased appetite?\n",
      "[[Choices]]\n",
      "TRUE\n",
      "FALSE\n\n",
      
      "[[Question: MC:SingleAnswer]]\n",
      "[[ID: case_", case_no, "_polyphagia]]\n",
      "Polyphagia?\n",
      "[[Choices]]\n",
      "TRUE\n",
      "FALSE\n\n",
      
      "[[Question: TE]]\n",
      "[[ID: case_", case_no, "_vomiting_reference]]\n",
      "<b>Vomiting:</b>\n<br><br>",
      "Reference: \n\n",
      
      "[[Question: MC:SingleAnswer]]\n",
      "[[ID: case_", case_no, "_vomiting]]\n",
      "Vomiting?\n",
      "[[Choices]]\n",
      "TRUE\n",
      "FALSE\n\n",
      
      "[[Question: TE]]\n",
      "[[ID: case_", case_no, "_defecation_reference]]\n",
      "<b>Defecation:</b>\n<br><br>",
      "Reference: \n\n",
      
      "[[Question: MC:SingleAnswer]]\n",
      "[[ID: case_", case_no, "_diarrhea]]\n",
      "Diarrhea?\n",
      "[[Choices]]\n",
      "TRUE\n",
      "FALSE\n\n",
      
      "[[Question: MC:SingleAnswer]]\n",
      "[[ID: case_", case_no, "_constipation]]\n",
      "Constipation?\n",
      "[[Choices]]\n",
      "TRUE\n",
      "FALSE\n\n",
      
      "[[Question: TE]]\n",
      "[[ID: case_", case_no, "_weight_reference]]\n",
      "<b>Weight:</b>\n<br><br>",
      "Reference: \n\n",
      
      "[[Question: MC:SingleAnswer]]\n",
      "[[ID: case_", case_no, "_weight_loss]]\n",
      "Weight loss?\n",
      "[[Choices]]\n",
      "TRUE\n",
      "FALSE\n\n"
    )
    survey_text <- paste0(survey_text, block_text)
  }
  
  return(survey_text)
}
```

## Make pilot survey
### Input
- pilot_hx_deidentified.xlsx (manually deidentified pilot hx set)
### Output
- qualtrics_survey_pilot_hx.txt (intermediates/make_survey)
### Issues / Deviations
Originally surveys were constructed and completeted in five batches a 20 records. Conglomerated here for simplicity (full script and files in LLM/vault)
The field "presenting complaint" was not included in the pilot survey (commented out)
```{r}
pilot_hx_deidentified <- read_xlsx(paste0(inputDir, "pilot_hx_deidentified.xlsx"))

survey_pilot_hx  <- generate_survey_text(pilot_hx_deidentified)

#writeLines(survey_pilot_hx, paste0(intermediatesDir,"survey_pilot_hx.txt"))
```

## Make tuning survey
### Input
- tuning_deid (intermediates/parse_deid/)
### Output
- survey_tuning.txt (intermediates/make_survey)
### Deviations/Issues
The field "presenting complaint" was not included in the tuning survey (commented out)
```{r make tuning survey}
tuning_deid <- fromJSON(paste0(toBox, "Projects/LLM/llm_vet_records/intermediates/parse_deid/tuning_deid.json"))

# Escape HTML tags (allows <> around redacted text to be visible in qualtrics survey)
tuning_deid <- tuning_deid %>%
  mutate(pertinentHistory_visit = sapply(pertinentHistory_visit, escape_html))

survey_tuning <- generate_survey_text(tuning_deid)

#writeLines(survey_tuning, paste0(intermediatesDir,"survey_tuning.txt"))
```

## Make test survey
### Input
- test_deid (intermediates/parse_deid/)
### Output
- survey_test.txt (intermediates/make_survey)
```{r make test survey}
# Escape html tags
test_deid <- fromJSON(paste0(toBox, "Projects/LLM/llm_vet_records/intermediates/parse_deid/test_deid.json")) %>%
  mutate(presentingComplaint_visit = sapply(presentingComplaint_visit, escape_html),
         pertinentHistory_visit = sapply(pertinentHistory_visit, escape_html))

survey_test <- generate_survey_text(test_deid)

# Save the survey text to a file
writeLines(survey_test, paste0(intermediatesDir,"survey_test.txt"))

```


