---
title: "parse_survey"
author: "Wulcan"
date: "2024-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objective
Parse qualtrics survey output and add mode and agreement if more then one respondent
```{r libraries, directories and functions}
library(tidyverse)
library(readxl)

toBox <- "~/Library/CloudStorage/Box-Box/"

inputDir <- paste0(toBox,"Projects/LLM/llm_vet_records/input/")

promptDir <- paste0(toBox,"Projects/LLM/llm_vet_records/prompts/")

intermediatesDir <- paste0(toBox,"Projects/LLM/llm_vet_records/intermediates/parse_survey/")

# Replace atypical quotation marks
clean_quotation_marks <- function(x) {
  str_replace_all(x, "[“”]", "\"")
}

parse_survey_response <- function(survey_response, records, prompt) {
  # Read and adjust qualtrics response
  survey_response <- survey_response %>%
    slice(-c(1, 2)) %>% 
  mutate(across(ends_with("_reference"), clean_quotation_marks)) %>%
    mutate(across(ends_with("_reference"), ~ ifelse(. == "", 'NA', .)))
  
  responseIds <- unique(survey_response$ResponseId)
  humans <- paste0("human_", seq_along(responseIds))
  name_mapping <- setNames(humans, responseIds)
  
 survey_response <- survey_response %>%
  mutate(respondent = recode(ResponseId, !!!name_mapping),
         respondent_type = "Humans",
         respondent_no = as.numeric(gsub("human_", "", respondent))) 

  # Select and reshape response columns
  response_long <- survey_response %>%
    pivot_longer(cols = starts_with("case_"),
                 names_to = c("case", ".value"),
                 names_pattern = "case_(\\d+)_(.*)") %>%
    select(respondent, respondent_type, respondent_no, case,
           appetite_reference, decreased_appetite, polyphagia, vomiting_reference, vomiting, defecation_reference, diarrhea, constipation, weight_reference, weight_loss) %>%
    mutate(case = as.integer(case)) %>%
    arrange(respondent, case)
  
  response_long$case <- as.numeric(response_long$case)
  
  logical_vars <- c("decreased_appetite", "polyphagia", "vomiting", "diarrhea", "constipation", "weight_loss")
  
  response_long[logical_vars] <- lapply(response_long[logical_vars], as.logical)

  # Summarize timing data per record
  timing_data <- survey_response %>%
    select(respondent, starts_with("case_")) %>%
    pivot_longer(cols = starts_with("case_"),
                 names_to = c("case", "variable"),
                 names_pattern = "case_(\\d+)_(.*)") %>%
    filter(variable == "timing_Page.Submit") %>%
    group_by(respondent, case) %>%
    summarise(elapsed_time = sum(as.numeric(value), na.rm = TRUE), .groups = 'drop')

  timing_data$case <- as.numeric(timing_data$case)

  # Combine the response with timing and other details
  response_parsed <- response_long %>%
    left_join(timing_data, by = c("respondent", "case")) %>%
    left_join(records, by = c("case" = "case")) %>%
    mutate(
           temperature = "Humans",
           prompt = prompt,
           record = paste(admission_date_visit, presentingComplaint_visit, pertinentHistory_visit, sep = " - "),
           respondent = respondent) %>% 
    mutate(across(c(vomiting, diarrhea, polyphagia, decreased_appetite, constipation, weight_loss), as.logical)) %>%
    select(respondent, respondent_type, respondent_no, case, appetite_reference, decreased_appetite, polyphagia, vomiting_reference, vomiting, defecation_reference, diarrhea, constipation, weight_reference, weight_loss, elapsed_time, prompt, record, temperature)
  

# Generate the long format
  response_parsed_long <- response_parsed %>%
    pivot_longer(
      cols = c(decreased_appetite, polyphagia, vomiting, diarrhea, constipation, weight_loss),
      names_to = "question",
      values_to = "response"
    ) %>%
    mutate(reference = case_when(
      question %in% c("decreased_appetite", "polyphagia") ~ appetite_reference,
      question == "vomiting" ~ vomiting_reference,
      question %in% c("diarrhea", "constipation") ~ defecation_reference,
      question == "weight_loss" ~ weight_reference
    )) %>% 
    select(-appetite_reference, -vomiting_reference, -defecation_reference, -weight_reference)

  # Generate the wide format
  response_parsed_wide <- response_parsed_long %>%
    pivot_wider(names_from = respondent, values_from = c(response, reference), names_glue = "{.value}_{respondent}")

# Check if there is more than one human respondent
  if (length(humans) > 1) {
    # Custom function to calculate the mode
    calculate_mode <- function(x) {
      uniq_x <- unique(x)
      uniq_x[which.max(tabulate(match(x, uniq_x)))]
    }

    # Add the mode column to the wide dataframe
    response_parsed_wide <- response_parsed_wide %>%
      rowwise() %>%
      mutate(mode = calculate_mode(c_across(starts_with("response_")))) %>%
      ungroup()

    # Function to calculate agreement score for each row
    calculate_agreement <- function(df) {
      # Get only the response columns
      response_cols <- df %>% select(starts_with("response_"))
      
      # Calculate pairwise agreement for each pair of observers
      pairwise_agreements <- combn(ncol(response_cols), 2, function(idx) {
        response_cols[[idx[1]]] == response_cols[[idx[2]]]
      })
      
      # Calculate the average agreement score
      avg_agreement <- rowMeans(pairwise_agreements, na.rm = TRUE)
      
      return(avg_agreement)
    }

    # Add the agreement column to the wide dataframe
    response_parsed_wide <- response_parsed_wide %>%
      mutate(agreement = calculate_agreement(.))
  }

  # Return all three versions as a list
  return(list(
    response_parsed = response_parsed,
    response_parsed_long = response_parsed_long,
    response_parsed_wide = response_parsed_wide
  ))
}
```

## Parse pilot survey
### Objective
Parse pilot survey from qualtrics 
### Input
- pilot_hx_response 
- pilot_hx_deidentified
- prompt_pilot_hx_text
### Output
pilot_hx_response_parsed (intermediates/parse_survey)

### Deviations
pilot_hx_deid doesn't include presenting complaint, I updated the parsing function after running this. Will fail if run again but to minor to adjust function
```{r parse pilot survey}
pilot_hx_response <- readRDS(paste0(inputDir, "pilot_hx_response.rds"))

pilot_hx_deidentified <- read_xlsx(paste0(inputDir, "pilot_hx_deidentified.xlsx"))

prompt_text <- readLines(paste0(promptDir, "prompt_pilot_hx.txt"), warn = FALSE) %>% paste(collapse = "\n")

results <- parse_survey_response(pilot_hx_response, pilot_hx_deidentified, prompt_text)

response_parsed <- results$response_parsed

#saveRDS(response_parsed, paste0(intermediatesDir, "pilot_hx_response_parsed.rds"))

```

## Parse tuning survey
### Objective
Parse tuning survey from qualtrics 
### Input
- tuning_response.csv
- tuning_deid.xlsx
- prompt_tuning_1.txt
### Output
tuning_response_parsed (intermediates/parse_survey)
tuning_response_parsed_wide (intermediates/parse_survey)

```{r parse tuning survey}
tuning_response <- read_csv(paste0(inputDir, "tuning_response.csv"))

tuning_deidentified <- read_xlsx(paste0(inputDir, "tuning_deid.xlsx"))

prompt_tuning_1 <- readLines(paste0(promptDir, "prompt_tuning_1.txt"), warn = FALSE) %>% paste(collapse = "\n")

results <- parse_survey_response(tuning_response, tuning_deidentified, prompt_tuning_1)

tuning_response_parsed <- results$response_parsed

saveRDS(tuning_response_parsed, paste0(intermediatesDir, "tuning_response_parsed.rds"))

tuning_response_parsed_wide <- results$response_parsed_wide

saveRDS(tuning_response_parsed_wide, paste0(intermediatesDir, "tuning_response_parsed_wide.rds"))
```

## Parse test survey
### Objective
Parse test survey from qualtrics 
### Input
- test_response.csv
- test_deid.xlsx
- prompt_test_1.txt
### Output
test_response_parsed (intermediates/parse_survey)
test_response_parsed_wide (intermediates/parse_survey)

```{r parse test survey}
test_response <- read_csv(paste0(inputDir, "test_response.csv"))

test_deidentified <- read_xlsx(paste0(inputDir, "test_deidentified.xlsx"))

prompt_test <- readLines(paste0(promptDir, "prompt_test.txt"), warn = FALSE) %>% paste(collapse = "\n")

results <- parse_survey_response(test_response, test_deidentified, prompt_test)

test_response_parsed <- results$response_parsed

saveRDS(test_response_parsed, paste0(intermediatesDir, "test_response_parsed.rds"))

test_response_parsed_wide <- results$response_parsed_wide

saveRDS(test_response_parsed_wide, paste0(intermediatesDir, "test_response_parsed_wide.rds"))
```