---
title: "parse_llm_output"
author: "Wulcan"
date: "2024-08-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objective
parse llm response
```{r libraries, directories and functions}
library(tidyverse)
library(jsonlite)

toBox <- "~/Library/CloudStorage/Box-Box/"
#toBox <- "C:/Users/15303/Box/"

inputDir <- paste0(toBox,"Projects/LLM/llm_vet_records/input/")

experimentDir <- paste0(toBox,"Projects/LLM/llm_vet_records/experiment/")

intermediatesDir <- paste0(toBox,"Projects/LLM/llm_vet_records/intermediates/parse_llm_output/")

# Define the function
parse_llm_response <- function(llm_response, model, output_token_price, input_token_price) {
  
  llm_response <- llm_response %>%
    rename(case = case_number,
           error = response) %>%
    mutate(cost_cents = ((num_output_tokens * output_token_price) + (num_input_tokens * input_token_price))/10)
  
  # Change name of and in response Id 
  responseIds <- unique(llm_response$iteration)
  llms <- paste0(model, "_", seq_along(responseIds))
  name_mapping <- setNames(llms, responseIds)
  
  llm_response <- llm_response %>%
    mutate(respondent = recode(iteration, !!!name_mapping),
           respondent_type = model,
           respondent_no = as.numeric(gsub(paste0(model, "_"), "", respondent))) %>%
    select(respondent, respondent_type, respondent_no, temperature, status, error, case, appetite_reference, 
           decreased_appetite, polyphagia, vomiting_reference, vomiting, 
           defecation_reference, diarrhea, constipation, weight_reference, 
           weight_loss, elapsed_time, num_output_tokens, num_input_tokens, cost_cents, prompt, record)
  
  # Prep temp var for merging
  llm_response$temperature <- as.character(llm_response$temperature)
  
  return(llm_response)
}
```

## Parse tuning GPT4o
### Input
- tuning_GPT4o (experiment/tuning_4o)
### Output
- tuning_GPT4o_parsed (intermediates/parse_llm_output)
```{r parse tuning GPT4o}
tuning_GPT4o <- fromJSON(paste0(experimentDir, "/tuning_4o/tuning_GPT4o.json"))

model_name <- "GPT4o"

output_token_price <- 0.015
  
input_token_price <- 0.005

tuning_GPT4o_parsed <- parse_llm_response(tuning_GPT4o, model_name, output_token_price, input_token_price)

saveRDS(tuning_GPT4o_parsed, paste0(intermediatesDir, "tuning_GPT4o_parsed.rds"))

```

## Parse test GPT4o
### Input
- test_GPT4o (experiment/test_4o)
### Output
- test_GPT4o_parsed (intermediates/parse_llm_output)
```{r test_GPT4o}
test_GPT4o <- fromJSON(paste0(experimentDir, "/test_4o/test_GPT4o.json"))

model_name <- "GPT4o"

output_token_price <- 0.015
  
input_token_price <- 0.005

test_GPT4o_parsed <- parse_llm_response(test_GPT4o, model_name, output_token_price, input_token_price)

saveRDS(test_GPT4o_parsed, paste0(intermediatesDir, "test_GPT4o_parsed.rds"))
```

## Parse test GPT4o corrected
### Input
- test_GPT4o_corrected (manually corrected json format)
### Output
- test_GPT4o_corrected_parsed (intermediates/parse_llm_output)
```{r test_GPT4o_corrected}
test_GPT4o_corrected <- fromJSON(paste0(inputDir, "test_GPT4o_corrected.json"))

model_name <- "GPT-4o"

output_token_price <- 0.015
  
input_token_price <- 0.005

test_GPT4o_corrected_parsed <- parse_llm_response(test_GPT4o_corrected, model_name, output_token_price, input_token_price)

saveRDS(test_GPT4o_corrected_parsed, paste0(intermediatesDir, "test_GPT4o_corrected_parsed.rds"))
```

## Parse test GPT-3.5
### Input
- test_GPT35 
### Output
- test_GPT35_parsed (intermediates/parse_llm_output)
```{r test GPT35}
test_GPT35 <- fromJSON(paste0(experimentDir, "/test_35/test_GPT35.json"))

model_name <- "GPT-3.5"

output_token_price <- 0.0015
input_token_price <- 0.0005

test_GPT35_parsed <- parse_llm_response(test_GPT35, model_name, output_token_price, input_token_price)

saveRDS(test_GPT35_parsed, paste0(intermediatesDir, "test_GPT35_parsed.rds"))
```