---
title: "sample_size"
author: "Wulcan"
date: "2024-06-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective
Calculate the sample size needed to estimate accuracy of LLM for extracting 
clinical information from veterinary records. 

### Study design
We are comparing a dichotomized index test (true/false LLM) to a dichotomized reference standard (true/false human majority opinion).

### Accuracy metrics
Sensitivity, Specicificy, positive predicitive value and negative predictive value.

### Sample size considerations
To estimate sample size for studies of the sensitivity/specificty of a single test,
Obuchowski (1998) provides a formula. For a condition with a prevalence below 50% 
sensitivity calculations will require a higher value then specificity.
In addition to deciding the acceptable margin of error and type I error rate,
the calculation requires an estimate of the sensitvity and prevalence of the condition.

Based on previous studies in humans we expect the sensitvity of the LLM to be ~ 95%
Based on the prevalence of the clinical signs of interests in a pilot set of 100 cases 
we expect the median prevalence of clinical signs investigated to be 15% 

A lower prevalence and/or sensitivity then expected will result in a wider confidence interval than intended. For low prevalence clinical signs (e.g. polyphagia and constipation we will probably have wide confidence intervals. Sizing the study after these would not be feasible)

### Sample size formula for constructing confidence intervals of length L:
```{r}
ci_length <- 0.07  #intended length of confidence interval
alpha <- 0.05 #type I error rate
sensitivity_estimate <- 0.95
prevalence_estimate <- 0.15

z_alpha_two_tail <- abs(qnorm(alpha_adjusted/2, lower.tail = TRUE))


variance_sensitvity_estimate <- sensitivity_estimate * (1-sensitivity_estimate) #binomial

n_true <- z_alpha_two_tail^2 * variance_sensitvity_estimate/ (ci_length^2) #the required number of trues

n_both <- n_true / prevalence_estimate #required total number (trues and falses)
```



