---
title: "sample_records"
author: "Wulcan"
date: "2024-08-14"
output: html_document
---

# Objective
A script to sample a set of records and provide some exploratory summary stats and exploration

# Issues and deviations
The current function does not automatically resample for duplicated (patientNo) but reports it. Will be adjusted in future projects but kept as is for reproducibility.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, directories and functions}
library(tidyverse)
library(writexl)

toBox <- "~/Library/CloudStorage/Box-Box/"
#toBox <- "C:/Users/15303/Box/" #lenovo

inputDir <- paste0(toBox,"Projects/LLM/llm_vet_records/input/")

intermediatesDir <- paste0(toBox,"Projects/LLM/llm_vet_records/intermediates/sample_records/")

outputDir <- paste0(toBox, "Projects/LLM/llm_vet_records/output/sample_records/")

# Function for pulling random sample (allowing exclusion of previously used)
  sample_records <- function(df_pool, used_dfs, n, seed) {
    
     # Check if used_dfs is empty
  if (length(used_dfs) == 0) {
    exclude <- character(0)  # Empty vector, meaning no exclusions
  } else {
    # Combine visitNos from all used_dfs
    exclude <- do.call(union, lapply(used_dfs, function(df) df$visitNo))
  }
  
  # Filter out the records that have already been used
  sampling_pool <- df_pool %>% filter(!visitNo %in% exclude)
  
  # Set the seed for reproducibility
  set.seed(seed)
  
  # Pull n random cases from the sampling pool
  sampled_records <- sampling_pool %>% sample_n(n)
  
  # Check for duplicate patientNos
  any_duplicates <- duplicated(sampled_records$patientNo) | duplicated(sampled_records$patientNo, fromLast = TRUE)
  
  # Print the result of the duplicate check
  if (any(any_duplicates)) {
    print("There are duplicates in the patientNo variable.")
  } else {
    print("There are no duplicates in the patientNo variable.")
  }
  
  # Get the duplicate values in patientNo
  duplicate_values <- unique(sampled_records$patientNo[duplicated(sampled_records$patientNo) | duplicated(sampled_records$patientNo, fromLast = TRUE)])
  
  # Remove duplicates
  removed_rows <- c()
  for (dup_value in duplicate_values) {
    dup_rows <- which(sampled_records$patientNo == dup_value)
    removed_row <- sample(dup_rows, 1)
    removed_rows <- c(removed_rows, removed_row)
    sampled_records <- sampled_records[-removed_row, ]
  }
  
  # Get remaining patientNos from the sampling pool
  remaining_patientNo <- setdiff(sampling_pool$patientNo, sampled_records$patientNo)
  
  # Add new records to replace removed duplicates
  for (i in seq_along(removed_rows)) {
    new_record <- sampling_pool %>%
      filter(patientNo %in% remaining_patientNo) %>%
      sample_n(1)
    
    sampled_records <- rbind(sampled_records, new_record)
    remaining_patientNo <- setdiff(remaining_patientNo, new_record$patientNo)
  }
  
  # Check again for duplicates
  any_duplicates <- duplicated(sampled_records$patientNo) | duplicated(sampled_records$patientNo, fromLast = TRUE)
  
  if (any(any_duplicates)) {
    print("There are duplicates in the patientNo variable.")
  } else {
    print("There are no duplicates in the patientNo variable.")
  }
  
  # Add a case number
  sampled_records <- sampled_records %>% mutate(case = row_number())
  
  # Generate the summary table
  summary_table <- sampled_records %>%
    summarise(
      year_min = min(year, na.rm = TRUE),
      year_max = max(year, na.rm = TRUE),
      year_mean = mean(year, na.rm = TRUE),
      year_median = median(year, na.rm = TRUE),
      year_sd = sd(year, na.rm = TRUE),
      year_25th = quantile(year, 0.25, na.rm = TRUE),
      year_75th = quantile(year, 0.75, na.rm = TRUE),
      wordCount_min = min(pertinentHistory_wordCount, na.rm = TRUE),
      wordCount_max = max(pertinentHistory_wordCount, na.rm = TRUE),
      wordCount_mean = mean(pertinentHistory_wordCount, na.rm = TRUE),
      wordCount_median = median(pertinentHistory_wordCount, na.rm = TRUE),
      wordCount_sd = sd(pertinentHistory_wordCount, na.rm = TRUE),
      wordCount_25th = quantile(pertinentHistory_wordCount, 0.25, na.rm = TRUE),
      wordCount_75th = quantile(pertinentHistory_wordCount, 0.75, na.rm = TRUE)
    )
  
  # Create histogram for 'year'
  histogram_year <- ggplot(sampled_records, aes(x = year)) +
    geom_histogram(binwidth = 1, fill = "#3282Bd", color = "black") +
    labs(x = "Year", y = "Frequency") +
    scale_x_continuous(limits = c(1991, 2023), breaks = seq(1991, 2023, by = 4)) +
    theme_bw() +
    theme(
      axis.text = element_text(size = 9, family = "Arial"),
      axis.title = element_text(size = 9, family = "Arial"),  
      legend.title = element_text(size = 9, family = "Arial")
    )
  
  # Create histogram for 'pertinentHistory_wordCount'
  histogram_words <- ggplot(sampled_records, aes(x = pertinentHistory_wordCount)) +
    geom_histogram(binwidth = 25, fill = "#3282Bd", color = "black") +
    labs(x = "Word Count", y = "Frequency") +
    theme_bw() +
    scale_x_continuous(limits = c(0, 1400), breaks = seq(0, 1400, by = 200)) + 
    theme(
      axis.text = element_text(size = 9, family = "Arial"),
      axis.title = element_text(size = 9, family = "Arial"),  
      legend.title = element_text(size = 9, family = "Arial")
    )
  
  # Return the sampled records, summary table, and plots
  list(
    sampled_records = sampled_records,
    summary_table = summary_table,
    histogram_year = histogram_year,
    histogram_words = histogram_words
  )
  }
  

#Function for selecting specific cases
select_cases <- function(cases, df) {
  # Filter the data to include only the selected cases
  selected_cases <- df %>%
    filter(case %in% cases)
  
  return(selected_cases)
}

```


## Sample pilot free text
### Objective
Generate a pilot set from EHRs (all free text fields)
### Input
- contains_free_text.rds <VMACS/all_cats_ever/output/contains_free_text>
### Output
- pilot_free_text.rds (also saved in vault for reproducibility)
- pilot_free_text_for_grading.xlsx (used for manual evaluations first pilot set)

```{r sample pilot free text}

#Read in all records with non-empty free text fields
contains_free_text <- readRDS(paste0(inputDir, "contains_free_text.rds"))

# Combine the used data frames in a list
used_dfs <- list()

# Call the function to get 270 new records
results <- sample_records(contains_free_text, used_dfs, n = 270, seed = 123)
pilot_free_text <- results$sampled_records

print(results$summary_table)
print(results$histogram_year)
print(results$histogram_words)

pilot_free_text <- results$sampled_records %>% 
  select(patientNo, visitNo, admission_date_visit, discharge_date_visit, age_visit, weight_visit, bcs_visit, presentingComplaint_visit, pertinentHistory_visit, physicalExamination_visit, problems_visit, procedures_visit, plans_and_progress_notes_visit, clinicalDx_visit, comments_visit, dischargeSummary_visit, dischargeInstructions_visit)

#saveRDS(pilot_free_text, paste0(intermediatesDir, "pilot_free_text.rds"))

#Create grading sheet with variables of interest
#discard everything expept visitNo
pilot_free_text_id <- pilot_free_text %>% select(visitNo)

#For each of the clinical signs, generate variables
clinical_signs <- c("lethargy", "decreased_appetite", "polyphagia", "vomiting", "diarrhea", "constipation", "weight_loss")

suffix_1 <- c("_presence", "_frequency_times_per_week", "_grade", "_acute_vs_chronic", "_duration_days")

suffix_2 <- "_citation"

# Loop through each clinical sign and each suffix to create new columns
for (sign in clinical_signs) {
  for (suf in suffix_1) {
    new_var <- paste0(sign, suf)
    pilot_free_text_id[[new_var]] <- "" # Initialize with empty string
    citation_var <- paste0(new_var, suffix_2)
    pilot_free_text_id[[citation_var]] <- "" # Initialize citation variable with empty string
  }
}

#remove variables that doesn;t make sense
pilot_free_text_id <- pilot_free_text_id %>% select(-lethargy_frequency_times_per_week,
                                  -lethargy_frequency_times_per_week_citation,
                                  -decreased_appetite_frequency_times_per_week, 
                                  -decreased_appetite_frequency_times_per_week_citation, - polyphagia_frequency_times_per_week, - polyphagia_frequency_times_per_week_citation, -polyphagia_grade, -polyphagia_grade_citation, -constipation_frequency_times_per_week, -constipation_frequency_times_per_week_citation, -weight_loss_frequency_times_per_week, -weight_loss_frequency_times_per_week_citation)

#save as excel
#write_xlsx(pilot_free_text_id, paste0(intermediatesDir,"pilot_free_text_for_grading.xlsx"))

```

## Sample pilot hx
### Objective
Generate a pilot set from EHRs (history fields only)
### Input
- contains_hx.rds (VMACS/all_cats_ever/output/contains_free_text/)
### Output
- pilot_hx.rds (also saved in vault for reproducibility)
- pilot_hx_for_deid.xlsx (template for manual deidentification)
### Issues / Deviations
Cases in pilot_free_text not excluded
```{r sample pilot hx}

#Read in all records with non-empty histories
#contains_hx <- readRDS(paste0(inputDir, "contains_hx.rds"))

# Combine the used data frames in a list
used_dfs <- list()

# Call the function to get 270 new records
results <- sample_records(contains_free_text, used_dfs, n = 100, seed = 234)
pilot_free_text <- results$sampled_records

print(results$summary_table)
print(results$histogram_year)
print(results$histogram_words)


#select_variables
pilot_hx_for_export <- pilot_hx %>% select(case, admission_date_visit, presentingComplaint_visit, pertinentHistory_visit) 

#Save rds
#saveRDS(pilot_hx, paste0(intermediates, "pilot_hx.rds"))
#write_xlsx(pilot_hx_for_export, paste0(intermediatesDir, "pilot_hx_for_deid.xlsx"))
```

## Select tuning set
### Objective
Generate a tuning set of handpicked EHRs from pilot_hx

### Input
- pilot_hx (intemediates/sample_records)

### Output
- tuning_for_deid (intermediates/sample_records)

```{r select tuning}


pilot_hx <- readRDS(paste0(intermediatesDir, "pilot_hx.rds")) %>% 
  select(case, admission_date_visit, presentingComplaint_visit, pertinentHistory_visit)

# Selected cases identified from pilot_hx as good tuning cases
selected_tuning <- c(30, 17, 10, 45, 49, 50, 51, 57, 78, 79)

tuning_cases <- select_cases(selected_tuning, pilot_hx)

#save as excel for manual removal of info
write_xlsx(tuning_cases, paste0(intermediatesDir, "tuning_for_deid.xlsx"))
```

## Select example
### Objective
Select a handpicked EHRs from pilot_hx for prompt example

### Input
- pilot_hx (intermediates/sample_records)

### Output
- example_for_deid (intermediates/sample_records)
```{r select example}
intermediatesDir <- paste0(toBox,"Projects/LLM/llm_vet_records/intermediates/sample_records/")

pilot_hx <- readRDS(paste0(intermediatesDir, "pilot_hx.rds")) %>% 
  select(case, admission_date_visit, presentingComplaint_visit, pertinentHistory_visit)

# Selected cases identified from pilot_hx as good tuning cases
selected_tuning <- 91

example_case  <- select_cases(selected_tuning, pilot_hx)

#save as excel for manual removal of info
write_xlsx(example_case, paste0(intermediatesDir, "example_for_deid.xlsx"))

```


## Sample test set
### Objective
Generate a test set from EHRs (history and presenting complaint fields)
### Input
- contains_hx.rds (VMACS/all_cats_ever/output/contains_free_text/)
- pilot_free_text.rds (intermediates/pilot)
- pilot_hx.rds (intermediates/pilot)
- pilot_hx_for_deid.xlsx (intermediates/pilot)

### Output
- test_250 (intermediates/test)
- test_250_for_deid (intermediates/test)
- year_words_test_set.csv (output/test)
```{r sample test}
#Read in all records with non-empty histories
contains_hx <- readRDS(paste0(inputDir, "contains_hx.rds"))

# Generate vector of visitNos to exclude (used in pilot sets)
pilot_free_text <- readRDS(paste0(intermediatesDir, "pilot_free_text.rds"))
pilot_hx <- readRDS(paste0(intermediatesDir, "pilot_hx.rds"))

# Combine the used data frames in a list
used_dfs <- list(pilot_free_text, pilot_hx)

# Call the function to get 250 new records
results <- sample_records(contains_hx, used_dfs, n = 250, seed = 345)
test_250 <- results$sampled_records

print(results$summary_table)
print(results$histogram_year)
print(results$histogram_words)

#saveRDS(test_250, paste0(intermediatesDir, "test_set.rds")) #hold on to for reproducibility

#select_variables for identification set
#test_250_for_deid <- test_250 %>% select(case, admission_date_visit, presentingComplaint_visit, pertinentHistory_visit)

#save as excel for manual removal of info
#write_xlsx(test_250_for_deid, paste0(intermediatesDir, "test_for_deid.xlsx"))

write_csv(results$summary_table, paste0(outputDir, "year_words_test_set.csv"))
```



